Stack Safety is a Security Property

### 2021-04-13 tex/main.tex
a few things in 3

The black rectangle represents the return address. In a typical RISC calling convention
this will have been stored in a register by call itself, and {\tt f} stores
it in the stack during its preamble, so it is not itself sealed.

@@

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       *($SP-1) = stash;
14:       stash = 0;
        }
15[R]:    return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l l}
      {\tt 2} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\goodc}[RetPtr]%
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1}
      \MemoryLabel{-15em}{0.75em}{3}
      \vspace{.5em} &
      \memory[3em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 3}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 10} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\goodc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{5} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
      \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{3} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{\bf 1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = 5} \\
    \end{tabular}
  \end{subfigure}
  \label{fig:controlflow}
  \caption{A Control Flow Attack}
\end{figure}

From the classic stack smashing attack to return oriented programming, 
one of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow is well-bracketed; rather,
they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''
in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to line 3 instead of 5; then it ends
up called a third time, at which point it returns to line 5.

Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
-1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
This has the appearance of an integrity violation. But how can that be?

The key intuition is that to return to {\tt main}, {\tt f} needs to do more than just
execute a return-labeled instruction. It must then actually reach the appropriate
{\em return target} in {\tt main}. In our context, we keep a stack of return targets,
which correspond to the depths at which stack frames have been sealed. A callee is
only considered to have returned when it executes a return-annotated instruction
and then reaches a valid return target. Only then do the sealed frames become unsealed.

>>>>

The black rectangle represents the return address. In a typical RISC calling convention
this will have been saved into a register by the call itself, and {\tt f} stores
it in the stack during its preamble, so it is not itself sealed.

@@

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       *($SP-1) = stash;
14:       stash = 0;
        }
15[R]:    return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l l}
      {\tt 2} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\goodc}[RetPtr]%
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1}
      \MemoryLabel{-15em}{0.75em}{\#3}
      \vspace{.5em} &
      \memory[3em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#3}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 10} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\goodc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{\#5} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{\#3} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{\bf 1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5} \\
    \end{tabular}
  \end{subfigure}
  \caption{A Control Flow Attack.}
  \label{fig:controlflow}
\end{figure}

%From the classic stack smashing attack to return oriented programming, 
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow is well-bracketed; rather,
they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}, a relative of the classic ``awkward example,''~\citep{??}
in which {\tt f} is called twice, at lines 2 and 4. It should therefore return to lines
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to line 3 instead of 5; then it ends
up being called a third time, at which point it returns to line 5.
\apt{I added \# marks to the line numbers to avoid confusion; this should
  probably be explained. What do the colors mean in this example?}

Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
-1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
This has the appearance of an integrity violation.  How do we capture this in our
integrity property?

The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs to do more than just
execute a return-labeled instruction: it must actually reach the appropriate
{\em return target} in {\tt main}. In our context\apt{what's that?}, we keep a stack of return targets,
which correspond to the depths at which stack frames have been sealed. A callee is
only considered to have returned when it executes a return-annotated instruction
and then reaches a valid return target. Only then do the sealed frames become unsealed.
